\documentclass[journal]{IEEEtran}

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{hyperref}
\usepackage{cite}
\usepackage[dvipsnames]{xcolor}
\usepackage{url}
\usepackage{array}
\usepackage{tabularx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{placeins}           % allow floats to drift across sections
\usepackage{float}              % allows [H] strict placement
\usepackage{tikz}               % native LaTeX diagrams (taxonomy figure)
\usetikzlibrary{positioning,arrows.meta,calc,fit,backgrounds}
% Tweak float parameters so pages fill more of their body before pushing floats
\renewcommand{\topfraction}{0.90}
\renewcommand{\bottomfraction}{0.75}
\renewcommand{\textfraction}{0.10}
\renewcommand{\floatpagefraction}{0.80}
\setcounter{topnumber}{3}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{5}

% A few helper macros
\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}}
\newcommand{\citeneeded}{\textcolor{red}{[CITE]}}

\title{Compression Techniques for Deep Stereo Matching}

\author{
\parbox{\linewidth}{\centering
Nazib~Abrar, Md~Raihanul~Haque~Rahi, and Md~Zunaid~Hossen \\[2pt]
\small Department of Mechatronics Engineering \\
Rajshahi University of Engineering \& Technology, Rajshahi, Bangladesh}%
\thanks{Manuscript prepared April 2026.}%
}

\begin{document}

\maketitle

\begin{abstract}
Deep stereo matching has progressed through three paradigms in under a decade: learned matching costs, end-to-end 3D cost-volume regression, and, most recently, the integration of monocular depth foundation models such as Depth Anything V2. The foundation-model era delivers state-of-the-art zero-shot accuracy, but at parameter counts and inference latencies that typically exceed embedded-accelerator budgets by an order of magnitude or more.

This survey catalogues the techniques that published methods have employed to deploy deep stereo matchers on edge hardware, organising them into seven compression families: backbone substitution, cost-volume compression, iterative-loop compression, knowledge distillation, architectural compression, adaptive compute, and neural architecture search. For each family we describe the underlying mechanism, review representative methods, and report the accuracy and latency trade-offs claimed in the original publications. We further review cross-domain generalisation techniques relevant to compressed models, and the limited but growing body of edge-hardware validation. The goal is descriptive: to map which compression strategies have been explored in the literature, how each compares to the accuracy envelope set by contemporary foundation-model stereo, and what aspects of edge deployment remain under-reported. We leave synthesis for a specific deployment scenario to the reader.
\end{abstract}

\begin{IEEEkeywords}
Stereo matching, deep learning, edge deployment, model compression, knowledge distillation, real-time inference, autonomous systems.
\end{IEEEkeywords}

\input{sections/01_introduction}
\input{sections/02_background}
\input{sections/03_foundation_baseline}
\input{sections/04_compression_taxonomy}
\input{sections/05_generalization}
\input{sections/06_edge_hardware}
\input{sections/07_roadmap}
\input{sections/08_conclusion}

\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}
